### 2023-07-26 paper/main.tex
added paper

%%
%% Copyright 2007-2020 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

%\documentclass[preprint,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
 \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{caption}
\usepackage{hyperref}

%% Hide "preprint submitted by" line
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

%% You might want to define your own abbreviated commands for common used terms, e.g.:
\newcommand{\kms}{km\,s$^{-1}$}
\newcommand{\msun}{$M_\odot}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={},
%%            city={},
%%            postcode={},
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{%
High-Throughput Phenotyping of Populus Trichocarpa Using Computer Vision \\
\large SMC Data Challenge \#1
}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[]{Vivaan Singhvi}
\author[]{Langalibelele Lunga}
\author[]{Pragya Nidhi}
\author[]{Chris Keum}
\author[]{Varrun Prakash}

\begin{abstract}
%% Text of abstract
	High-throughput phenotyping refers to the non-destructive and efficient evaluation of plant phenotypes. Machine Learning methods offer the prospect of addressing two major challenges in phenotyping plants: efficiency in handling large datasets and extraction of specific traits. Previous techniques, such as the application of Deep Neural Nets in tandem with automated cameras, to our best knowledge, refrain from using physically labeled plants to organize datasets. To accomplish this, we used Optical Character Recognition(OCR) to read the physical labels on the plants, image segmentation techniques(OpenCv, FacebookSAM)  in order to apply Machine Learning algorithms(RandomForestClassifier, XGradientBoosting) for plant morphology classifications, and analyzed encoded EXIF tag information for the purpose of finding correlations between phenotypes. We used a  Populus Trichocarpa  plant dataset provided by Oak Ridge National Laboratory, with 1,672 image files of plants labeled with treatment( Control or Drought), block, row, position, and genotype. We found that OCR techniques, such as PaddleOCR, successfully read 91.3\% of the labeled plants, allowing for labeled information to be accurately placed in a pandas Dataframe. The success of the OCR carried into our XGradientBoosting and RandomForestClassifier techniques, which classified leaf morpholgoies by leaf shape and color with an accuracy of 62.82\%. Finally, we found [insert findings from step 4]. The use of Machine Learning in High-throughoutPut Phenotyping has shown to be effective in analyzing large plant datasets[Improve this sentence]. It allows for efficient and environmentally friendly classification of plants.
\end{abstract}

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
%\begin{highlights}
%\item Research highlight 1
%\item Research highlight 2
%\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword, up to a maximum of 6 keywords
computer vision \sep machine learning \sep high-throughput phenotyping

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%\tableofcontents

%% \linenumbers

%% main text

\section{Introduction}
\label{intro}

Image-based phenotyping refers to a breakthrough technology used in plant biology and agriculture to examine and assess plants’ “anatomical, ontological, physiological, and biochemical features” through the use of images. Previous studies have shown its potential as a noninvasive replacement for traditional on field techniques used to extract important phenotypic data. In recent years, this potential has been further intensified by coupling image-based phenotyping techniques with machine learning algorithms. This new approach, called high throughput phenotyping, has enabled the extraction of phenotypes from “complex” plant image datasets that were previously challenging to analyze efficiently. Furthermore, more recent studies have shown that the technology holds potential in identifying correlations between phenotype, genotype, and environmental metadata.

In this study, we aim to develop an accurate Machine Learning Model for High Throughput Phenotyping of leaves’ morphological traits (e.g. leaf size, shape, color) in plant images; unlike datasets in previous research, the dataset used in this study contains physical labels within the images, each containing important information on treatment (Control or Drought), block, row, position, and genotype. As a result, this study will deal with a new aspect of image-based phenotyping namely, extracting data from the white labels to identify correlations between leaves’ phenotypes and data embedded within the white labels. Given this information, the study aims to answer the following questions:

\begin{enumerate}

\item Is it possible to use optical character recognition (OCR) or machine learning techniques to “Read” the label on each tag and generate a spreadsheet containing the treatment, block, row, position, and genotype? Doing this would dramatically simplify data collection, as this information is usually collected manually.

\item Can machine learning differentiate and classify different leaf morphologies among genotypes by classifying leaf shape or color characteristics?

\item Can a predictive model be built using leaf morphology classifications that may indicate that a particular genotype was cultivated in a “drought” or “control” condition?

\item GPS and other camera information are encoded in EXIF tags. Can this data be used to determine characteristics such as leaf size? Can other data, such as soil maps, weather, etc. be used to find correlations among phenotypes?

\end{enumerate}

\section{Related Works}
\label{related_works}

Traditionally, plant morphologies would have to be analyzed by hand, wasting time and potentially damaging the environment. Now, with breakthroughs in high throughput image phenotyping led by Machine Learning techniques, methods such as Automated Machine Learning \citep{koh2021automated}, Computer Vision with Deep Learning \citep{mochida2019computer},  and Convolutional Neural Networks(CNNs) \citep{pound2017deep}, have allowed for efficient and timely analysis of plant phenotypes. With these methods, researchers are able to extract features from plants and classify plant morphologies effectively with little effect on the plant environment.

\subsection{Advancements in High Throughput Phenotyping Techniques}
\label{related_works_1}

Advancements in the field of plant phenotyping have been pioneered by technology. 
From using hardware, such as Raspberry Pi imaging systems \citep{tausen2020greenotyper} to thermographical sensing techniques \citep{walter2015plant}, research in the field of high phenotyping has led to an acceleration in working efficiency. Plants are now able to be analyzed in fast, effective, and accurate methods allowing for plant to be used in scientific discoveries. In our study, we rely on machine learning techniques to classify morphologies based on a phone image dataset provided.

\subsection{Machine Learning Approaches for Phenotypic Analysis}
\label{related_works_2}

Machine Learning has enabled researchers with computer vision, allowing them to create programs that “read” image information in real time. In addition, it increases the efficiency of obtaining leaf morphology data. However, researchers have found that the use of traditional classification techniques, such as RandomForrestClassifier, have a high performance, but do not generalize well across datasets \citep{pound2017deep}. In order to increase accuracy of models and the generalization of the model, Deep Learning techniques have been introduced. With deep learning models, models train themselves iteratively until they reach a desired outcome. These techniques, such as Convolutional Neural Networks \citep{koh2021automated, pound2017deep} have shown to be highly accurate in classifying plant structures while maintaining performance across different datasets. Convolutional Networks consist of several layers, which allow for more discriminative and detailed analysis of the plant image, leading to highly accurate classifications. In our study, we use traditional methods(GradientBoosting, RandomForest) to classify plant morphology, as our focus is on organizing the plant images into an easily readable dataset and performing classifications, instead of creating models highly transferable across datasets.

\subsection{Applications of High Throughput Phenotyping with Machine Learning}
\label{related_works_3}

Due to global events, such as climate change and global population increase, the ability to produce large amounts of healthy food will be crucial to society. With Machine Learning being used for phenotyping, researchers will be able to rapidly analyze food crops to maximize production and crop breeding \citep{arya2022deep, shakoor2017high}.
As climate change and global population increases, global crop yield will have to increase to provide for the growing populations. Further research has shown that climate change will continue to negatively affect crops and cause crop diseases \citep{newton2011implications}, hence the need for effective and precise analysis of plants, which is done with Machine Learning models. In our study, we aim to show how Machine Learning can be used to organize plant data, dissect plant images into meaningful classifications, and allow for rapid investigation of plants for scientists in the field of plant phenotyping.

\section{Methodology}
\label{method}
%%\label{}

Given the nature of the project, there are many aspects requiring obtaining large files, such as the dataset itself or large models. Therefore, all scripts needed to fully set up the project (downloading large files or setting up a large virtual environment), are viewable in the \verb|scripts/| folder in the GitHub repository, linked in Section \ref{code}.

All solutions to the challenge were implemented in Python 3.9/3.10. Several notable libraries used throughout the project include: OpenCV \citep{opencv_library}, the open-source computer vision library; Scikit-Learn \citep{scikit-learn}, Python implementations of dozens of machine learning algorithms and data processing tools; Pandas \citep{mckinney2010data}, a data manipulation library featuring the powerful \verb|DataFrame| object; and NumPy \citep{harris2020array}, an array manipulation library crucial for fast operations on images.

\subsection{Reading Labels with Optical Character Recognition}
\label{step_1}

To read the text, a pre-trained Optical Character Recognition (OCR) model seemed like the optimal choice. However, there are several models high-performing models available for use. The three candidates chosen for this project were \verb|PyTesseract| \citep{lee2017pytesseract}, \verb|EasyOCR| \citep{jaided2020easyocr}, and \verb|PaddleOCR| \citep{du2020ppocr}.

The three models were tested using metrics of performance and efficiency, which were measured through their accuracy score and time taken on a sample dataset of nine images. The results of the testing can be viewed in \ref{app_ocr_testing}.

\verb|PaddleOCR| was superior in speed, being significantly faster than \verb|EasyOCR| and marginally faster than \verb|PyTesseract|. It is also more accurate in both measurements, having a higher accuracy score with and without null values.

In order to fix the aforementioned null values from the OCRs, images went through augmentation, with subsequent attempts being made to read text during each step. First, images are rotated 45 degrees in order to fix potential orientation issues. Then, the original image is thresholded using OpenCV's \verb|adaptiveThreshold| to amplify the edges. Finally, if nothing works, the image is both rotated and thresholded.

\begin{center}
    \includegraphics[width=0.45\textwidth]{images/step1_pipeline.png}
    \captionof{figure}{Different versions of images tried for optical character recognition, in order}
\end{center}

After the text was read by the OCR, regular expressions (RegEx) were used to find the precise details mentioned in the dataset: treatment, block, row, position, and genotype. A subjective analysis of the images in the dataset showed that the treatment was either a ‘C’ or a ‘D’, and the remaining features followed the RegEx patterns below (note: \verb|\d| represents a digit, anything between 0-9) :

\begin{verbatim}
  Block:      B\d+
  Row:        R\d+
  Position:   P\d+
  Genotype:   [A-Z]{2,}(-\d+)+(_\d+)*(_[A-Z]+)?
\end{verbatim}

After all the text is read and processed, it is converted to a Pandas \verb|DataFrame| and saved to an Excel file.

\subsection{Classifying Leaf Morphologies with Image Processing}
\label{step_2}

In order to be able to determine the morphological characteristics of the plants, it is necessary to perfectly segment the leaves from each image. This proved to be a challenging endeavor due to the complexity of the background of the images, other datasets, such as the Komatsuna dataset, \citep{uchiyama2017easy} had 'cleaner' image backgrounds.

To obtain perfect masks for leaves in each image, the Segment Anything Model (SAM) \citep{kirillov2023segment} is the obvious choice. Being a relatively new and groundbreaking model, this paper is one of the first to utilize a pre-trained model of its caliber.

The \verb|AutomaticMaskGenerator| provided generates accurate masks for everything in an image. However, it has two issues: it may generate too many masks (especially for images as cluttered as those in the dataset), and is computationally expensive, taking too long to be a plausible approach for all 1672 images in this dataset. Through a \verb|SamPredictor| in conjunction with an ONNX \citep{bai2019onnx} \verb|InferenceSession|, masks would generate much quicker, and only from desired points. However, to obtain these desired points, the images would have to be processed to approximate the locations of the leaves.

The first step was to hide non-leaf elements in the image. A popular technique called HSV filtering was employed, used in a similar project by \cite{szachowicz2021komatsuna}. This involves filtering all pixels that are not between an HSV-encoded color range, helping eliminate many background objects such as wood, dirt, and of course, the label.

Then, OpenCV's implementation of the \verb|Canny| algorithm was used to highlight significant edges in the image, which were present around the boundary of leaves but were minimal inside of them. However, some leaves had boundary edges with gaps, causing most contours detected (using OpenCV's \verb|findContours|) to be sporadic. Therefore, the edges were dilated in multiple iterations by using a large kernel, causing the boundaries to be closed. After contours were found on the resulting image, they were filtered by pruning those that had a height or width (found using the \verb|boundingRect| function around each contour) too small compared to that of the image. Additionally, contours with not enough green inside them (namely less than $\frac{100}{255}$ green) were removed.

\begin{center}
	\centering
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/1.png}	
        \includegraphics[width=0.15\textwidth]{images/seg_pipeline/2.png}      	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/3.png}	
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/4.png}	
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/5.png}	
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/6.png}	
	\captionof{figure}{Pipeline of image transformations for leaf approximation, read left to right}
	% \label{fig_mom0}%
\end{center}

To find the approximate points of these leaf contours, the midpoint was calculated using this simple formula, with $x, y, w, h$ being the outputs of OpenCV's \verb|boundingRect| function:

\begin{equation}
    \mbox{midpoint} = (x + \frac{w}{2}, y + \frac{h}{2})
\end{equation}

By using these points as target points for the SAM/ONNX mask predictor, leaves could reliably be obtained from the image. Masks could then be saved for each image by converting each of an image's masks to a random uniform grayscale value and combining them all into an image. While very accurate, however, some leaves detected by the program were not appropriate for use in morphology analysis. For example, some leaves could be cut off by another object, or, in rare cases, be green wood misinterpreted as leaves.

Thus, it was necessary to train a machine learning model that was able to accurately detect these kinds of 'bad' leaves. Training data was generated by showing randomly selected leaves and getting human input regarding their suitability, extracting their features, and storing the data in a file. The model was implemented using Scikit-Learn's \verb|RandomForestClassifier|. The model had an accuracy of 90.91\% on testing data generated by Scikit-Learn's \verb|train_test_split|.

\begin{center}
	\centering
	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/6.png}	
        \includegraphics[width=0.15\textwidth]{images/seg_filter/1.png}      	\includegraphics[width=0.15\textwidth]{images/seg_filter/2.png}	
	\captionof{figure}{The process for filtering segmentations through the model: from contour approximations to generated masks to filtered masks}
	% \label{fig_mom0}%
\end{center}

Then, each leaf could be cropped and rotated, which, when combined with masking every pixel outside the its border, effectively highlighted and isolated it for visual purposes. By repeating this process for each image in the dataset, random leaves could then be selected to generate data for training the morphological classification model.

By using a slightly modified version of the aforementioned feature extraction function in conjunction with human labeling of leaves regarding their shape, color, and level of brown splotches, a \verb|MultiOutputClassifier| was able to be implemented with Scikit-Learn. More specifically, the \verb|XGBoost| algorithm was used for each of the three predicted features in the \verb|MultiOutputClassifier|.

TO BE CONTINUED (REWRITING NOT FINISHED YET).

\subsection{Treatment Predictions using Morphological Classifications}
\label{step_3}

The task for this step was to build a model to predict if a leaf was grown in a control or drought environment, given morphological classifications from the previous step.

The challenge for using classification data as features to train a model with is that they are categorical, rather than quantitative variables; thus, extra transformations must be done to properly represent the data.

To properly represent the classifications, the method of one-hot encoding (MAYBE FIND SOURCE SAYING ITS GOOD) was implemented using the Pandas \verb|get_dummies| method. This is a process in which a feature is split into multiple columns, with each row having a one if it matches the column or a zero if it doesn't.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{c||c c c c} 
 \hline
 leaf\_color & light\_green & dark\_green & yellow\_green & yellow  \\
 \hline
 light\_green & 1 & 0 & 0 & 0 \\
 yellow\_green & 0 & 0 & 1 & 0 \\
 dark\_green & 0 & 1 & 0 & 0 \\ 
 light\_green & 1 & 0 & 0 & 0 \\
 yellow & 0 & 0 & 0 & 1 \\
 \hline
\end{tabular}
}
\captionof{table}{Example transformation of a column with one-hot encoding, with the original column on the left and the encoded columns on the right}
\label{Table1}
\end{center}

This process was done with the leaf color and shape variables. However, since the level of brown splotches is ordinal, it was assigned levels 0, 1, 2, and 3, corresponding to ‘none’, ‘low’, ‘medium,’ and ‘high.’ Now, with all the data numerical, it could be used to train and test the model.

\subsection{Using EXIF Data to Assess Leaf Size and Analyze Correlations Between Morphologies and Environments}
\label{step_4}

Firstly, we aimed to measure the size of a leaf through metadata, encoded in EXIF Tags, which are embedded within each of the images. These EXIF tags typically give important information about the images, including the type of camera utilized, the distance of the camera from a photo, and photo dimensions (in pixels). In order to initiate the extraction of such metadata from the images, a package called “exif” was used and gave a variety of retrieval options.

Out of all the metadata extracted from the EXIF tags, the most notable pieces of information included ResolutionX/Y tags and focal length. Although these pieces of information are crucial in developing a mechanism to measure leaf size in the dataset, the information provided simply does not suffice in determining leaf size; we needed information such as how far the leaf was from the camera, information that was not given from EXIF tags alone. This lack of information rendered the metadata as inadequate for direct measurement of leaf size.

Another approach we considered to measure leaf size included using the white labels, held up by the researcher in each image, as a known reference object. Although this method seemed promising, there were certain limitations in most images in the dataset that prevented a consistent measurement. Firstly, the white label in the image was often obstructed by the researcher’s hands, as shown in Figure 1. This obstruction would ultimately interfere with alignment of the white label, making the label unreliable as an object of reference. Another limitation to this method came from the fact that the labels were often tilted, not in the same plane as the leaves. The tilt of the labels caused inaccuracies in measurement and would lead to misrepresentation of the true length. Along with that, there was no way of finding out how far forward or behind the leaf was relative to the tag, meaning the method would result in an inaccurate representation of the leaf.

In the second part of challenge 4, we aimed to find correlation among phenotypes using geographical features including soil type and weather conditions. Using the EXIF tags, we were able to retrieve the coordinates where the image was taken. We noticed all of the images contained were taken in approximately the same area, meaning weather would be the same for all plants. Along with that, most soil map APIs contain information about the soil in an area, so it would be unable to distinguish miniscule changes in locations. For all other potential geographical changes, due to the close proximity all images are taken in, we were unable to find out if there were any correlations. Not only was the location the same, the images were taken in a short time frame, within a week. This gave us the same limitations as the close location had.

\section{Results}
%%\label{}

\subsection{Reading Labels with Optical Character Recognition}
\label{step_1}

\subsection{Classifying Leaf Morphologies with Image Processing}
\label{step_2}

\subsection{Treatment Predictions using Morphological Classifications}
\label{step_3}

\subsection{Using EXIF Data to Assess Leaf Size and Analyze Correlations Between Morphologies and Environments}
\label{step_4}

Overall, the metadata embedded in these specific EXIF tags did not prove to be enough for direct measurement of leaf size. Despite two alternative approaches being used, there were always absences in crucial information. Future studies could include additional calibration features to enable the development of a more feasible method for leaf size assessment. For the second part of the challenge, we needed larger differences in location or times that the images were taken in order to study how the conditions in those periods affected the phenotypes.

\section{Conclusions and Significance}

High throughput phenotyping has gained prominence due to its potential in solving a wide variety of agricultural problems. With the world population projected to reach 9.3 billion people by 2050 and a need to produce 60\% more food \citep{silva2012feeding}, it will become crucial for researchers to discover a method for identifying productive genotypes for plant breeding.  With machine learning methods, such as Optical Character Recognition (OCR), classifiers such as RandomForest, and image segmentation techniques,  productivity in the field of plant/crop production will increase rapidly, along with the accurate analysis of plants.

By successfully applying an OCR to read labels, our research shows potential for automating a data extraction process.  This process will significantly reduce the manual labor needed to organize information and allow for larger datasets to be handled with ease.

In addition, the use of XGradientBoosting and the RandomForestClassifier for the classification of plant phenotype allows for precise examination of plant morphologies. Although the achieved accuracy of our RandomForestClassifier, 62.82\%, shows room for improvement, it highlights the potential the method holds in classification. Future optimization may enable higher accuracy whilst still providing a feasible method for non-invasively extracting valuable information about plant morphologies from onsite images.

Finally, the analysis of EXIF tags provided by the dataset shows promise in estimating the dimensions of an image as well as identifying correlations between plant phenotypes. Although the EXIF tags did not prove sufficient for these specific tasks, we have outlined a few crucial data points that may be helpful/necessary in order to assess dimensions and identify correlations in future studies.

\section{Code Availability}
\label{code}

All code is publicly available under the MIT License on GitHub here: \url{https://github.com/vivaansinghvi07/smoky-mountain-data-comp}.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix

\section{Results of OCR Model Testing}
\label{app_ocr_testing}

\begin{center}
	\centering
	\includegraphics[width=0.4\textwidth]{images/ocr_average_accuracy.png}	
	\captionof{figure}{Average accuracy score for reading the sample dataset for the three OCR models}
	% \label{fig_mom0}%
\end{center}

\begin{center}
	\centering
	\includegraphics[width=0.4\textwidth]{images/pruned_ocr_average_accuracy.png}	
	\captionof{figure}{Average accuracy score only including non-zero values (only images that were able to be read)}
	% \label{fig_mom0}%
\end{center}

\begin{center}
	\centering
	\includegraphics[width=0.4\textwidth]{images/ocr_time_taken.png}	
	\captionof{figure}{Total time taken for the three OCR models on the sample dataset}
	% \label{fig_mom0}%
\end{center}

%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%https://www.overleaf.com/project/64bec4b7b47d12a55db700f5

\bibliographystyle{elsarticle-harv}
\bibliography{main}
%% else use the following coding to input the bibitems directly in the
%% TeX file.

%%\begin{thebibliography}{00}

%% \bibitem[Author(year)]{label}
%% For example:

%% \bibitem[Aladro et al.(2015)]{Aladro15} Aladro, R., Martín, S., Riquelme, D., et al. 2015, \aas, 579, A101

%%\end{thebibliography}

\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.

### 2023-07-26 scripts/build_paper.sh
added apper building script

@@ -1,0 +1,57 @@
+#!/bin/bash
+
+red=$(printf "\033[31m")
+green=$(printf "\033[32m")
+reset=$(printf "\033[0m")
+
+# check if user on mac
+if [ $(uname) == "Darwin" ]
+then
+
+    # lead user to the installer 
+    if [ ! $(which port) ]
+    then
+        echo "You need to install MacPorts."
+        exit
+    fi
+
+    # installs brew if needed
+    if [ ! $(which brew) ]
+    then
+        /bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"
+    fi
+
+    # install dependencies
+    if [ ! $(which pdflatex) ]
+    then
+        brew install basictex
+    fi
+
+    # get things that may or may not be needed for template 
+    sudo port install texlive-math-science
+    sudo port install texlive-publishers
+
+else 
+
+    # install dependencies on Linux
+    sudo apt-get update 
+    sudo apt-get install texlive-latex-base
+    sudo apt-get install texlive-math-science
+    sudo apt-get install texlive-publishers
+
+fi
+
+# run script to create the paper
+cd paper
+pdflatex main
+bibtex main
+pdflatex main 
+pdflatex main
+
+# remove unneeded files
+rm main.aux
+rm main.log
+rm main.spl
+rm main.bbl
+rm main.blg
+rm main.out

### 2023-07-26 scripts/build_paper.sh
updated script

@@ -41,8 +41,10 @@ else
 
 fi
 
-# run script to create the paper
+# holds all files needed
 cd paper
+
+# run script to create the paper
 pdflatex main
 bibtex main
 pdflatex main 
@@ -55,3 +57,6 @@ rm main.spl
 rm main.bbl
 rm main.blg
 rm main.out
+
+# cd out of the paper dir
+cd ..

### 2023-07-26 scripts/startup.sh
adjusted startup

@@ -1,6 +1,6 @@
 # run basic startup scripts already written
 . scripts/venv.sh
-bash scripts/load_files.sh
+./scripts/load_files.sh
 
 # get the files required for step 2
 cd step2
@@ -10,3 +10,5 @@ then
 fi
 python3 load_onnx_model.py
 cd ..
+
+./scripts/build_paper.sh

### 2023-07-26 scripts/startup.sh
code look nicer

@@ -1,6 +1,7 @@
 # run basic startup scripts already written
 . scripts/venv.sh
 ./scripts/load_files.sh
+./scripts/build_paper.sh
 
 # get the files required for step 2
 cd step2
@@ -10,5 +11,3 @@ then
 fi
 python3 load_onnx_model.py
 cd ..
-
-./scripts/build_paper.sh

### 2023-07-26 scripts/load_files.sh
broken shebang

@@ -1,4 +1,4 @@
-#!bin/bash
+#!/bin/bash
 
 if [ ! $(which unzip) ]
 then

### 2023-07-28 paper/main.tex
final draft

@@
-%%
-%% Copyright 2007-2020 Elsevier Ltd
-%%
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%%
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%%
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
-
-%\documentclass[preprint,12pt,authoryear]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times,authoryear]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,authoryear]{elsarticle}
 \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
 
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
+\usepackage{caption}
+\usepackage[hidelinks]{hyperref}
 
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-\usepackage{caption}
-\usepackage{hyperref}
-
-%% Hide "preprint submitted by" line
 \makeatletter
@@
 \makeatother
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
-
-%% You might want to define your own abbreviated commands for common used terms, e.g.:
-\newcommand{\kms}{km\,s$^{-1}$}
-\newcommand{\msun}{$M_\odot}
 
@@
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%            addressline={},
-%%            city={},
-%%            postcode={},
-%%            state={},
-%%            country={}}
-%% \fntext[label3]{}
-
-\title{%
-High-Throughput Phenotyping of Populus Trichocarpa Using Computer Vision \\
+\title{
+High-Throughput Phenotyping using Computer Vision and Machine Learning \\
 \large SMC Data Challenge \#1
@@
 
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-
 \author[]{Vivaan Singhvi}
-\author[]{Langalibelele Lunga}
+\author[]{Langalibalele Lunga}
 \author[]{Pragya Nidhi}
@@
 High-throughput phenotyping refers to the non-destructive and efficient evaluation of plant phenotypes.
-Machine Learning methods offer the prospect of addressing two major challenges in phenotyping plants: efficiency in handling large datasets and extraction of specific traits.
-Previous techniques, such as the application of Deep Neural Nets in tandem with automated cameras, to our best knowledge, refrain from using physically labeled plants to organize datasets.
-To accomplish this, we used Optical Character Recognition(OCR) to read the physical labels on the plants, image segmentation techniques(OpenCv, FacebookSAM)  in order to apply Machine Learning algorithms(RandomForestClassifier, XGradientBoosting) for plant morphology classifications, and analyzed encoded EXIF tag information for the purpose of finding correlations between phenotypes.
-We used a  Populus Trichocarpa  plant dataset provided by Oak Ridge National Laboratory, with 1,672 image files of plants labeled with treatment( Control or Drought), block, row, position, and genotype.
-We found that OCR techniques, such as PaddleOCR, successfully read 91.3\% of the labeled plants, allowing for labeled information to be accurately placed in a pandas Dataframe.
-The success of the OCR carried into our XGradientBoosting and RandomForestClassifier techniques, which classified leaf morpholgoies by leaf shape and color with an accuracy of 62.82\%.
-Finally, we found [insert findings from step 4].
-The use of Machine Learning in High-throughoutPut Phenotyping has shown to be effective in analyzing large plant datasets[Improve this sentence].
-It allows for efficient and environmentally friendly classification of plants.
+In recent years, it has been coupled with machine learning in order to improve the process of phenotyping plants by increasing efficiency in handling large datasets and developing methods for the extraction of specific traits.
+Previous studies have developed methods to advance these challenges through the application of deep neural networks in tandem with automated cameras; however, the datasets being studied often excluded physical labels.
+In this study, we used a dataset provided by Oak Ridge National Laboratory with 1,672 images of Populus Trichocarpa with white labels displaying treatment (control or drought), block, row, position, and genotype.
+Optical character recognition (OCR) was used to read these labels on the plants, image segmentation techniques in conjunction with machine learning algorithms were used for morphological classifications, machine learning models were used to predict treatment based on those classifications, and analyzed encoded EXIF tags were used for the purpose of finding leaf size and correlations between phenotypes.
+We found that our OCR model successfully read 91.3\% of the labeled plants, allowing for the information to be accurately placed in a spreadsheet.
+Our classification models identified leaf shape, color, and level of brown splotches with an average accuracy of 62.82\%, and plant treatment with an accuracy of 60.08\%.
+Finally, we identified a few crucial pieces of information absent from the EXIF tags that prevented the assessment of the leaf size.
+There was also missing information that prevented the assessment of correlations between phenotypes and conditions.
+However, future studies could improve upon this to allow for the assessment of these features.
+The use of machine learning and computer vision in high-throughput phenotyping has shown to be effective in analyzing large plant datasets, leading to a more comprehensive phenotype analysis in plants and showing potential in various agricultural and environmental applications.
 \end{abstract}
 
-%%Graphical abstract
-%\begin{graphicalabstract}
-%\includegraphics{grabs}
-%\end{graphicalabstract}
-
-%%Research highlights
-%\begin{highlights}
-%\item Research highlight 1
-%\item Research highlight 2
-%\end{highlights}
-
 \begin{keyword}
-%% keywords here, in the form: keyword \sep keyword, up to a maximum of 6 keywords
 computer vision \sep machine learning \sep high-throughput phenotyping
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
 \end{keyword}
@@
 
-Image-based phenotyping refers to a breakthrough technology used in plant biology and agriculture to examine and assess plants’ “anatomical, ontological, physiological, and biochemical features” through the use of images.
-Previous studies have shown its potential as a noninvasive replacement for traditional on field techniques used to extract important phenotypic data.
+\subsection{Background Information}
+\label{background}
+
+High-throughput phenotyping is defined as a breakthrough technology used in plant biology and agriculture to examine and assess plants' "anatomical, ontological, physiological, and biochemical features" through the use of images.
+Previous studies have shown its potential as a noninvasive replacement for traditional on-field techniques used to extract important phenotypic data.
 In recent years, this potential has been further intensified by coupling image-based phenotyping techniques with machine learning algorithms.
-This new approach, called high throughput phenotyping, has enabled the extraction of phenotypes from “complex” plant image datasets that were previously challenging to analyze efficiently.
-Furthermore, more recent studies have shown that the technology holds potential in identifying correlations between phenotype, genotype, and environmental metadata.
+This new approach has enabled the extraction of phenotypes from "complex" plant image datasets that were previously challenging to analyze efficiently.
+Furthermore, newer studies have shown that the technology has potential to identifying correlations between phenotype, genotype, and environmental metadata.
 
-In this study, we aim to develop an accurate Machine Learning Model for High Throughput Phenotyping of leaves’ morphological traits (e.g.
-leaf size, shape, color) in plant images; unlike datasets in previous research, the dataset used in this study contains physical labels within the images, each containing important information on treatment (Control or Drought), block, row, position, and genotype.
-As a result, this study will deal with a new aspect of image-based phenotyping namely, extracting data from the white labels to identify correlations between leaves’ phenotypes and data embedded within the white labels.
-Given this information, the study aims to answer the following questions:
+\subsection{Research Objective}
+\label{objective}
+
+In this study, we aim to develop an accurate machine learning model for high-throughput phenotyping of leaves’ morphological traits (e.g.
+leaf size, shape, color) in plant images; unlike datasets in previous research, the dataset used in this study contains physical labels within the images, each containing important information on treatment (control or drought), block, row, position, and genotype.
+As a result, this study will address a new aspect of image-based phenotyping, namely, extracting data from the white labels to identify correlations between leaves’ phenotypes and data embedded within the white labels.
+During this project, we aim to answer the following questions:
 
@@
 
-\item Is it possible to use optical character recognition (OCR) or machine learning techniques to “Read” the label on each tag and generate a spreadsheet containing the treatment, block, row, position, and genotype?
+\item Is it possible to use optical character recognition (OCR) or machine learning techniques to “read” the label on each tag and generate a spreadsheet containing the treatment, block, row, position, and genotype?
 Doing this would dramatically simplify data collection, as this information is usually collected manually.
@@
 
-Traditionally, plant morphologies would have to be analyzed by hand, wasting time and potentially damaging the environment.
-Now, with breakthroughs in high throughput image phenotyping led by Machine Learning techniques, methods such as Automated Machine Learning \citep{koh2021automated}, Computer Vision with Deep Learning \citep{mochida2019computer},  and Convolutional Neural Networks(CNNs) \citep{pound2017deep}, have allowed for efficient and timely analysis of plant phenotypes.
-With these methods, researchers are able to extract features from plants and classify plant morphologies effectively with little effect on the plant environment.
+Traditionally, plant morphologies would be analyzed manually, wasting time and potentially damaging the environment.
+However, current breakthroughs in high-throughput phenotyping led by machine learning techniques\citep{koh2021automated, pound2017deep} and computer vision with deep learning \citep{mochida2019computer} have allowed for efficient and timely analyses of plant phenotypes.
+With these methods, researchers can extract features from plants and classify their morphologies effectively while having little effect on the plant itself.
 
@@
 Advancements in the field of plant phenotyping have been pioneered by technology.
-From using hardware, such as Raspberry Pi imaging systems \citep{tausen2020greenotyper} to thermographical sensing techniques \citep{walter2015plant}, research in the field of high phenotyping has led to an acceleration in working efficiency.
-Plants are now able to be analyzed in fast, effective, and accurate methods allowing for plant to be used in scientific discoveries.
+From using hardware, such as Raspberry Pi imaging systems \citep{tausen2020greenotyper} to thermographic sensing techniques \citep{walter2015plant}, research in the field of high-throughput phenotyping has led to an acceleration in working efficiency.
+Plants are now able to be analyzed in fast, effective, and accurate methods allowing for them to be used in scientific discoveries.
 In our study, we rely on machine learning techniques to classify morphologies based on a phone image dataset provided.
@@
 
-Machine Learning has enabled researchers with computer vision, allowing them to create programs that “read” image information in real time.
+Machine learning has enabled researchers with computer vision, allowing them to create programs that “read” image information in real-time.
 In addition, it increases the efficiency of obtaining leaf morphology data.
-However, researchers have found that the use of traditional classification techniques, such as RandomForrestClassifier, have a high performance, but do not generalize well across datasets \citep{pound2017deep}.
-In order to increase accuracy of models and the generalization of the model, Deep Learning techniques have been introduced.
+However, researchers have found that traditional classification techniques, such as Random Forest Classification, have high performance but do not generalize well across datasets \citep{pound2017deep}.
+In order to increase the accuracy of models and the generalization of the model, deep learning techniques have been introduced.
 With deep learning models, models train themselves iteratively until they reach a desired outcome.
-These techniques, such as Convolutional Neural Networks \citep{koh2021automated, pound2017deep} have shown to be highly accurate in classifying plant structures while maintaining performance across different datasets.
-Convolutional Networks consist of several layers, which allow for more discriminative and detailed analysis of the plant image, leading to highly accurate classifications.
-In our study, we use traditional methods(GradientBoosting, RandomForest) to classify plant morphology, as our focus is on organizing the plant images into an easily readable dataset and performing classifications, instead of creating models highly transferable across datasets.
+These techniques, such as convolutional neural networks \citep{koh2021automated, pound2017deep} have shown to be highly accurate in classifying plant structures while maintaining performance across different datasets.
+Convolutional neural networks consist of several layers, which allow for a more discriminative and detailed analysis of plant images, leading to highly accurate classifications.
+In our study, we use traditional methods to classify plant morphology, as our focus is performing classifications with a single dataset rather than creating models highly transferable across datasets.
 
@@
 
-Due to global events, such as climate change and global population increase, the ability to produce large amounts of healthy food will be crucial to society.
-With Machine Learning being used for phenotyping, researchers will be able to rapidly analyze food crops to maximize production and crop breeding \citep{arya2022deep, shakoor2017high}.
-As climate change and global population increases, global crop yield will have to increase to provide for the growing populations.
-Further research has shown that climate change will continue to negatively affect crops and cause crop diseases \citep{newton2011implications}, hence the need for effective and precise analysis of plants, which is done with Machine Learning models.
-In our study, we aim to show how Machine Learning can be used to organize plant data, dissect plant images into meaningful classifications, and allow for rapid investigation of plants for scientists in the field of plant phenotyping.
+Due to global events, such as climate change and global population increase, the ability to produce large amounts of healthy crops will be crucial to society.
+With machine learning being used for phenotyping, researchers will be able to rapidly analyze food crops to maximize production and breeding \citep{arya2022deep, shakoor2017high}.
+
+As climate change and the global population increase, global crop yield will have to increase to provide for the growing populations.
+Further research has shown that climate change will continue to negatively affect crops and cause crop diseases \citep{newton2011implications}, hence the need for effective and precise analysis of plants, which can be done with machine learning models.
+In our study, we aim to show how machine learning can be used to organize plant data, dissect plant images into meaningful classifications, and allow for rapid investigation of plants for scientists in the field of plant phenotyping.
 
@@
 Given the nature of the project, there are many aspects requiring obtaining large files, such as the dataset itself or large models.
-Therefore, all scripts needed to fully set up the project (downloading large files or setting up a large virtual environment), are viewable in the \verb|scripts/| folder in the GitHub repository, linked in Section \ref{code}.
+Therefore, all scripts needed to fully set up the project (downloading large files or setting up a large virtual environment), are viewable in the \verb|scripts/| folder of the GitHub repository, linked in Section \ref{code}.
 
@@
 Several notable libraries used throughout the project include: OpenCV \citep{opencv_library}, the open-source computer vision library; Scikit-Learn \citep{scikit-learn}, Python implementations of dozens of machine learning algorithms and data processing tools; Pandas \citep{mckinney2010data}, a data manipulation library featuring the powerful \verb|DataFrame| object; and NumPy \citep{harris2020array}, an array manipulation library crucial for fast operations on images.
+
+All models used in this project are trained and tested on data split by Scikit-Learn's \verb|train_test_split|.
+Additionally, Scikit-Learn's \verb|accuracy_score| is used to determine the accuracy of said models.
 
@@
 To read the text, a pre-trained Optical Character Recognition (OCR) model seemed like the optimal choice.
-However, there are several models high-performing models available for use.
+However, there are several high-performing models available for use.
 The three candidates chosen for this project were \verb|PyTesseract| \citep{lee2017pytesseract}, \verb|EasyOCR| \citep{jaided2020easyocr}, and \verb|PaddleOCR| \citep{du2020ppocr}.
@@
 After the text was read by the OCR, regular expressions (RegEx) were used to find the precise details mentioned in the dataset: treatment, block, row, position, and genotype.
-A subjective analysis of the images in the dataset showed that the treatment was either a ‘C’ or a ‘D’, and the remaining features followed the RegEx patterns below (note: \verb|\d| represents a digit, anything between 0-9) :
+A subjective analysis of the images in the dataset showed that the treatment was either a \verb|C| or a \verb|D|, and the remaining features followed the RegEx patterns below (note: \verb|\d| represents a digit, anything between 0-9) :
 
@@
 
-After all the text is read and processed, it is converted to a Pandas \verb|DataFrame| and saved to an Excel file.
+After all the text is read and processed, it is converted to a Pandas \verb|DataFrame| and saved to an Excel spreadsheet, which will be used in the following steps.
+Then, 30 random images are selected and the accuracy of the results is tested by comparing inputted to read features.
+As with the OCR testing earlier, it is calculated both with and without unread values.
 
@@
 
-In order to be able to determine the morphological characteristics of the plants, it is necessary to perfectly segment the leaves from each image.
-This proved to be a challenging endeavor due to the complexity of the background of the images, other datasets, such as the Komatsuna dataset, \citep{uchiyama2017easy} had 'cleaner' image backgrounds.
+In order to be able to determine the morphological characteristics of the plants, it was necessary to perfectly segment the leaves from each image.
+This proved to be a challenging endeavor due to the complexity of the background of the images; other datasets, such as the Komatsuna dataset, \citep{uchiyama2017easy} had 'cleaner' image backgrounds.
 
-To obtain perfect masks for leaves in each image, the Segment Anything Model (SAM) \citep{kirillov2023segment} is the obvious choice.
-Being a relatively new and groundbreaking model, this paper is one of the first to utilize a pre-trained model of its caliber.
+To obtain perfect masks for leaves in each image, the Segment Anything Model (SAM) \citep{kirillov2023segment} was the obvious choice, due to its groundbreaking accuracy and ease of use.
+Being a relatively new model, this paper is one of the first to utilize a pre-trained model of its caliber.
 
@@
 	\centering
-	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/1.png}	
-        \includegraphics[width=0.15\textwidth]{images/seg_pipeline/2.png}      	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/3.png}	
-	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/4.png}	
-	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/5.png}	
-	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/6.png}	
-	\captionof{figure}{Pipeline of image transformations for leaf approximation, read left to right}
+	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/1.png}	
+        \includegraphics[width=0.23\textwidth]{images/seg_pipeline/2.png}      	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/3.png}	
+	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/4.png}	
+	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/5.png}	
+	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/6.png}	
+	\captionof{figure}{Pipeline of image transformations for leaf approximation, read left to right. The last image visualizes detected contours}
 	% \label{fig_mom0}%
@@
 Training data was generated by showing randomly selected leaves and getting human input regarding their suitability, extracting their features, and storing the data in a file.
-The model was implemented using Scikit-Learn's \verb|RandomForestClassifier|.
-The model had an accuracy of 90.91\% on testing data generated by Scikit-Learn's \verb|train_test_split|.
+The model was implemented using Scikit-Learn's \verb|RandomForestClassifier|, and had an accuracy of 90.91\% on testing data.
 
@@
 	\centering
-	\includegraphics[width=0.15\textwidth]{images/seg_pipeline/6.png}	
-        \includegraphics[width=0.15\textwidth]{images/seg_filter/1.png}      	\includegraphics[width=0.15\textwidth]{images/seg_filter/2.png}	
-	\captionof{figure}{The process for filtering segmentations through the model: from contour approximations to generated masks to filtered masks}
+	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/6.png}	
+        \includegraphics[width=0.23\textwidth]{images/seg_filter/2.png}	
+	\captionof{figure}{From leaf approximations with image processing to filtered leaf segmentations, from the SAM/ONNX Runtime (green leaves fit for classification)}
 	% \label{fig_mom0}%
@@
 
-Then, each leaf could be cropped and rotated, which, when combined with masking every pixel outside the its border, effectively highlighted and isolated it for visual purposes.
+Then, each leaf could be cropped and rotated, which, when combined with masking every pixel outside its border, effectively highlighted and isolated it for visual purposes.
 By repeating this process for each image in the dataset, random leaves could then be selected to generate data for training the morphological classification model.
 
-By using a slightly modified version of the aforementioned feature extraction function in conjunction with human labeling of leaves regarding their shape, color, and level of brown splotches, a \verb|MultiOutputClassifier| was able to be implemented with Scikit-Learn.
-More specifically, the \verb|XGBoost| algorithm was used for each of the three predicted features in the \verb|MultiOutputClassifier|.
+The features selected for morphology prediction were: color, between light-green, dark-green, yellow-green, and yellow; shape, between ovate, lanceolate, elliptical, and oblong \citep{nakano2020plant}; and level of brown splotches (representing the level of the withering of the leaf), between none, low, medium, and high.
 
-TO BE CONTINUED (REWRITING NOT FINISHED YET).
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/leaf_shapes.jpeg}
+    \captionof{figure}{Leaf shape reference from the article by \cite{nakano2020plant}}
+\end{center}
+
+By using a slightly modified version of the aforementioned feature extraction function in conjunction with 
+a Scikit-Learn \verb|LabelEncoder| to prepare the target features, the classifier could be implemented with Scikit-Learn's \verb|MultiOutputClassifier|.
+This is a framework for using the same algorithm on three different features at once; namely, the algorithm used was \verb|XGBoost|, which seemed to trump other models in terms of accuracy.
+
+After the model was trained, it was used on each image in the dataset by finding all eligible leaves and determining the mode classification after each eligible leaf was run through the aforementioned model.
+The results were saved in a copy of the spreadsheet from the previous step.
 
@@
 
-The challenge for using classification data as features to train a model with is that they are categorical, rather than quantitative variables; thus, extra transformations must be done to properly represent the data.
+The classifications were easily attainable because they were saved in a spreadsheet.
+However, the challenge for using classification data as features to train a model with is that they are categorical, rather than quantitative; data needed to undergo extra transformations to prepare it for use in predictions.
 
-To properly represent the classifications, the method of one-hot encoding (MAYBE FIND SOURCE SAYING ITS GOOD) was implemented using the Pandas \verb|get_dummies| method.
+To properly represent the classifications, the method of one-hot encoding \citep{brownlee2020why} was implemented using the Pandas \verb|get_dummies| method.
 This is a process in which a feature is split into multiple columns, with each row having a one if it matches the column or a zero if it doesn't.
@@
 \resizebox{\columnwidth}{!}{%
-\begin{tabular}{c||c c c c} 
+\begin{tabular}{|c||c c c c|} 
  \hline
@@
 This process was done with the leaf color and shape variables.
-However, since the level of brown splotches is ordinal, it was assigned levels 0, 1, 2, and 3, corresponding to ‘none’, ‘low’, ‘medium,’ and ‘high.’ Now, with all the data numerical, it could be used to train and test the model.
+However, since a leaf's level of brown splotches is ordinal, it was assigned levels 0, 1, 2, and 3, corresponding to 'none', 'low', 'medium,' and ‘high.' The data for expected treatments were obtained by reading rows of the spreadsheet that had treatments successfully read.
+
+Now, using this data, a predictor could be trained and deployed.
+The chosen classifier for this task was Scikit-Learn's \verb|RandomForestClassifier|, as it has shown to be a capable model throughout the project.
+After the model was trained, it was deployed on the spreadsheet, predicting treatments for rows that lacked a properly read one.
 
@@
 
-Firstly, we aimed to measure the size of a leaf through metadata, encoded in EXIF Tags, which are embedded within each of the images.
-These EXIF tags typically give important information about the images, including the type of camera utilized, the distance of the camera from a photo, and photo dimensions (in pixels).
-In order to initiate the extraction of such metadata from the images, a package called “exif” was used and gave a variety of retrieval options.
+First, we aimed to measure the size of a leaf through metadata encoded in EXIF Tags, which are embedded within each of the images.
+These tags typically give important information about the images, including the type of camera utilized, the distance of the camera from a photo, and photo dimensions (in pixels).
+In order to initiate the extraction of such metadata from the images, the \verb|exif| package was used and gave a variety of retrieval options.
 
-Out of all the metadata extracted from the EXIF tags, the most notable pieces of information included ResolutionX/Y tags and focal length.
-Although these pieces of information are crucial in developing a mechanism to measure leaf size in the dataset, the information provided simply does not suffice in determining leaf size; we needed information such as how far the leaf was from the camera, information that was not given from EXIF tags alone.
-This lack of information rendered the metadata as inadequate for direct measurement of leaf size.
+Out of all the metadata extracted from the EXIF tags, the most notable pieces of information included \verb|ResolutionX|/\verb|ResolutionY| tags and \verb|FocalLength|.
+Although these pieces of information can assist in developing a mechanism to measure leaf size in the dataset, it simply does not suffice alone; additional information, such as how far the leaf was from the camera, was needed.
+With more information, such as \verb|FocalPlaneResolution| - showing the number of pixels per real unit (\verb|in| or \verb|cm|) - in addition to previously mentioned features, it is possible to assess the size of specific components of an image.
+For example, a study on maize phenotyping by \cite{liu2021pocketmaize} showed that embedded information in images was enough for a smartphone application to accurately determine measurements of plant height and leaf area in a photo.
 
@@
 Although this method seemed promising, there were certain limitations in most images in the dataset that prevented a consistent measurement.
-Firstly, the white label in the image was often obstructed by the researcher’s hands, as shown in Figure 1.
-This obstruction would ultimately interfere with alignment of the white label, making the label unreliable as an object of reference.
+First, the white label in the image was often obstructed by the researcher’s hands, as shown in \ref{app_blocked}.
+This obstruction would ultimately interfere with the alignment of the white label, making the label unreliable as an object of reference.
 Another limitation to this method came from the fact that the labels were often tilted, not in the same plane as the leaves.
@@
 Along with that, there was no way of finding out how far forward or behind the leaf was relative to the tag, meaning the method would result in an inaccurate representation of the leaf.
+Again, this issue could have been resolved with more careful documentation of images without obstruction and a better-positioned label alignment.
 
-In the second part of challenge 4, we aimed to find correlation among phenotypes using geographical features including soil type and weather conditions.
+In the second part of challenge 4, we aimed to find a correlation between phenotypes and environmental features, such as soil type and weather conditions.
 Using the EXIF tags, we were able to retrieve the coordinates where the image was taken.
-We noticed all of the images contained were taken in approximately the same area, meaning weather would be the same for all plants.
-Along with that, most soil map APIs contain information about the soil in an area, so it would be unable to distinguish miniscule changes in locations.
-For all other potential geographical changes, due to the close proximity all images are taken in, we were unable to find out if there were any correlations.
-Not only was the location the same, the images were taken in a short time frame, within a week.
-This gave us the same limitations as the close location had.
+We noticed all of the images contained were taken in approximately the same two locations, so any weather or soil API would be unable to provide specific conditions.
+If we were given this information as part of the dataset, we would have been better able to find correlations.
+Regardless, a study on how the temperature affects the phenotype and development plasticity in Arabidopsis Thaliana \citep{ibanez2017ambient} placed the plants in different temperature environments in order to measure their differences.
+If we had more information regarding the conditions in which the plants were raised, we would be better able to determine if there was a correlation between phenotypes and environment.
 
-\section{Results}
-%%\label{}
+\section{Results and Discussion}
+\label{results}
+
+All confusion matrices used in this section originate from Scikit-Learn's \verb|confusion_matrix| and \verb|ConfusionMatrixDisplay|.
 
@@
 
+The OCR was able to read 91.3\% of the labels in the images of the entire dataset.
+On the 30 test images, the model had a 77.33\% accuracy on extracted features (94.31\% with null values omitted).
+A subjective analysis showed that for all features except for genotype, the interpreted values seem to match the true values almost always.
+
+\begin{center}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{|c|c|c|c|c|c} 
+ \hline
+ filename & treatment & block & row & position & genotype \\
+ \hline
+    ... & D & 1 & 8  & 32 & BESC-34 \\
+    ... & C & 1 & 10 & 12 & **BESC-417\_LM**,core \\
+    ... & C & 2 & 3  & 40 & BESC-468 \\
+    ... & C & 2 & 6  & 54 & BESC-28\_LM \\
+    ... & C & 1 & 24 & 22 & **LILD-26-5\_LM**,core \\
+ \hline
+\end{tabular}
+}
+\captionof{table}{First five rows of data saved to the spreadsheet from step 1, with filenames omitted to conserve space}
+\label{Table1}
+\end{center}
+
+Before the text was inserted into the above spreadsheet, it was present in a Pandas \verb|DataFrame|.
+Here, it was possible to analyze the statistics of the data using the \verb|info| method, containing useful information such as null value count:
+
+\begin{verbatim}
+    RangeIndex: 1672 entries, 0 to 1671
+    Data columns (total 6 columns):
+     #   Column     Non-Null Count  Dtype  
+    ---  ------     --------------  -----  
+     0   filename   1672 non-null   object 
+     1   treatment  1098 non-null   object 
+     2   block      1306 non-null   float64
+     3   row        1388 non-null   float64
+     4   position   1414 non-null   float64
+     5   genotype   1431 non-null   object 
+\end{verbatim}
+
+Evidently, the former attributes were read the least compared to those near the end of the tag, with 66\% non-null values for the treatment and 86\% non-null values for the genotype.
+This is likely due to the many forms of obstructions in the images, such as a leaf, a data collector's hand, or even the boundary of the image itself.
+Examples of such can be seen in \ref{app_blocked}.
+
+Even though the information suggests that there is more error in reading the treatment, it is more likely that it stems from text processing.
+The interpreted genotype has more room for error, as there is an extensive number of possibilities for it, while the treatment must be a \verb|C| or a \verb|D|, leaving little room for error.
+
+Regardless, it is undeniable that the use of \verb|PaddleOCR|, a relatively new and underused model, proved significantly beneficial and successful in reading labels from the images in the dataset.
+However, one evident issue lies in the fact that the suffix of the genotype (such as \verb|_LM|) was often not read properly, causing the text processing to drop it entirely.
+Increased text pre-processing may have alleviated this issue.
+
 \subsection{Classifying Leaf Morphologies with Image Processing}
 \label{step_2}
+
+The spreadsheet from the last step was successfully updated with morphological predictions from this step.
+
+\begin{center}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{c|c|c|c|c|} 
+ \hline
+ genotype & leaf\_color & leaf\_shape & brown\_splotches \\
+ \hline
+    BESC-34	&                  light\_green &   ovate &	    none     \\
+    **BESC-417\_LM**,core &	   yellow\_green	&  ovate &      high     \\
+    BESC-468 &	               yellow &        lanceolate &	high     \\
+    BESC-28\_LM &	               dark\_green &	   ovate &      low      \\
+    **LILD-26-5\_LM**,core &	   yellow\_green	&  lanceolate &	medium   \\
+ \hline
+\end{tabular}
+}
+\captionof{table}{A continuation of Table 2, with morphological data added from step 2}
+\label{Table1}
+\end{center}
+
+The model had an average accuracy of 62.82\%, which, when considering the fact that each feature had four different classes, is a reasonable accuracy level.
+The confusion matrices generated by comparing test and predicted data for each feature are below:
+
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/morph_conf_matrix/color.png}
+    \captionof{figure}{Confusion matrix for testing data on the leaf color predictor}
+\end{center}
+
+This part of the model had an accuracy score of 69.23\%.
+It seems to be reliable in predicting leaf color but has difficulty classifying true yellow-green leaves; however, it still correctly predicts yellow-green most of the time.
+Because the testing data seems to lack yellow leaves, the model's accuracy regarding them is unknown.
+
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/morph_conf_matrix/shape.png}
+    \captionof{figure}{Confusion matrix for testing data on the leaf shape predictor}
+\end{center}
+
+While mostly reliable with color, the model is sporadic in predicting leaf shape, with an accuracy of 50.00\%.
+When leaves are truly shaped as oblong or ovate, the predictions are essentially random, including many more wrong guesses than correct guesses.
+However, this may be partially due to the subtlety of the difference between leaf shapes, which may be difficult for a model to predict.
+
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/morph_conf_matrix/splotch.png}
+    \captionof{figure}{Confusion matrix for testing data on the brown splotch level predictor }
+\end{center}
+
+While there is more 'noise' in the predictions for brown splotch levels when compared to those for leaf color, the model seems to have a moderate degree of accuracy (at 69.23\%), with the exception of classifying leaves with "medium" splotch levels.
+
+A source for much of the error in the model's predictions may lie in human error.
+For example, when labeling data, we may have incorrectly labeled any of the three features for a given leaf due to their aforementioned ambiguity.
+It is also worth noting that many of the classes were difficult to distinguish even subjectively, such as distinguishing between yellow-green and light-green colors or between elliptical and ovate leaf shapes.
 
@@
 
+The predictive model on the treatment of the plants had an accuracy of 60.08\%.
+
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/treat_conf_matrix.png}
+    \captionof{figure}{Confusion matrix for the treatment prediction model}
+\end{center}
+
+While more predictions were true than false, it is undeniable that the model is unreliable for determinate results.
+The inaccuracy is likely due to several factors, including the fact that the only features used in the model were the morphological classifications from step 2, which may have been inaccurate to begin with.
+Additionally, there may have been overfitting on training data.
+
 \subsection{Using EXIF Data to Assess Leaf Size and Analyze Correlations Between Morphologies and Environments}
@@
 
-Overall, the metadata embedded in these specific EXIF tags did not prove to be enough for direct measurement of leaf size.
-Despite two alternative approaches being used, there were always absences in crucial information.
+Overall, the metadata embedded in the EXIF tags specific to the dataset did not prove to be enough for direct measurement of leaf size.
+Given more useful tags, such as \verb|FocalPlaneResolution|, the task may have been possible.
+Additionally, using the label as a reference for measurement also proved impossible due to the various augmentations of the labels in images.
 Future studies could include additional calibration features to enable the development of a more feasible method for leaf size assessment.
-For the second part of the challenge, we needed larger differences in location or times that the images were taken in order to study how the conditions in those periods affected the phenotypes.
+For the second part of the challenge, we needed larger differences in locations or times that the images were taken in order to study how the conditions in those periods affected the phenotypes.
 
-\section{Conclusions and Significance}
+\section{Conclusions}
 
@@
 With the world population projected to reach 9.3 billion people by 2050 and a need to produce 60\% more food \citep{silva2012feeding}, it will become crucial for researchers to discover a method for identifying productive genotypes for plant breeding.
-With machine learning methods, such as Optical Character Recognition (OCR), classifiers such as RandomForest, and image segmentation techniques,  productivity in the field of plant/crop production will increase rapidly, along with the accurate analysis of plants.
+With computer vision techniques, such as Optical Character Recognition and image processing, and machine learning methods, such as classifiers to predict morphologies and treatments, productivity in the field of plant/crop production will increase rapidly, along with the accurate analysis of plants.
 
-By successfully applying an OCR to read labels, our research shows potential for automating a data extraction process.
-This process will significantly reduce the manual labor needed to organize information and allow for larger datasets to be handled with ease.
+\subsection{Originality and Uniqueness}
 
-In addition, the use of XGradientBoosting and the RandomForestClassifier for the classification of plant phenotype allows for precise examination of plant morphologies.
-Although the achieved accuracy of our RandomForestClassifier, 62.82\%, shows room for improvement, it highlights the potential the method holds in classification.
-Future optimization may enable higher accuracy whilst still providing a feasible method for non-invasively extracting valuable information about plant morphologies from onsite images.
+This project is one of the first to utilize \verb|PaddleOCR| \citep{du2020ppocr}, an extremely lightweight and capable Optical Character Recognition model, in the context of reading plant labels.
+Unlike other approaches, it is resilient to augmentations in data, such as label rotation or minor interruptions in labels.
 
-Finally, the analysis of EXIF tags provided by the dataset shows promise in estimating the dimensions of an image as well as identifying correlations between plant phenotypes.
-Although the EXIF tags did not prove sufficient for these specific tasks, we have outlined a few crucial data points that may be helpful/necessary in order to assess dimensions and identify correlations in future studies.
+It is also one of the first to use the Segment Anything Model (SAM) \citep{kirillov2023segment} in the context of leaf segmentation.
+When combined with image processing, leaves can now be extracted from an image with no training or processing steps for the user.
+The SAM allows for much more accurate masks to be obtained and is robust towards heavily noisy data such as ours.
+This is especially important when leaf shape needs to be classified, as accurate masks have more accurate features.
+
+Through the use of pre-built classifiers like \verb|RandomForestClassifier|, this paper shows that such models can be useful despite the little effort used by the programmer.
+
+\subsection{Significance and Specific Contributions}
+
+By successfully applying an OCR to read labels, our research shows potential for automating a data extraction process, which can significantly reduce the manual labor needed to organize information.
+Large datasets can be processed within minutes, with minimal effort from the user.
+
+The use of machine learning on leaves allows for easy analysis of plant morphologies, which reduces the need for subjective determinations.
+Although the model has room for improvement, the process shows promise and can be easily optimized.
+Moreover, using another model to predict the kind of environment a plant was grown, as was done in step 3, allows for even more valuable information to be obtained just from a single photograph.
+Researchers can save hours of effort analyzing thousands of images by utilizing tools such as these.
+
+Even though the EXIF tags provided by the dataset proved insufficient for determining leaf size, with the proper additional information, this task could become trivial and greatly benefit researchers who would otherwise have to manually measure each leaf with a ruler.
+
+\subsection{Limitations and Possible Improvements}
+
+To achieve higher accuracy in OCR label reading, future research could use datasets that are more uniform with clearer labels.
+On a similar note, the images in the dataset could have a clear focus on a specific leaf, making leaf classification models better.
+
+The dataset's lack of metadata prevented the use of EXIF tags to determine leaf size and correlations between phenotypes and environmental features.
+Therefore, a dataset with more information both embedded into images and embedded into separate files (such as files with soil condition information) could allow researchers to meet these objectives.
+
+However, if the same dataset is used, future researchers could build more accurate and complex neural networks for classification.
+The models used in this paper were all prebuilt and not fine-tuned to our needs.
+Additionally, the image processing could have improved reliability, making data more fit for use in training or feature prediction.
+
+\subsection{Future Directions}
+
+Future research could be done in expanding the species analyzed for morphology classifications.
+In this study, only Populus Trichocarpae were used, so the models used may only be effective on this species.
+
+Since a pretrained segmentation model (the SAM) was used in this study, researchers could attempt to build segmentation models fine-tuned to only recognize leaves, which could increase model efficiency and provide more consistent results.
 
@@
 
+\section*{Acknowledgements}
+\label{acknowledgements}
+
+We would like to thank Dr. Ty Frazier for his suggestions and mentoring throughout this project.
+
 %% The Appendices part is started with the command \appendix;
 %% appendix sections are then done as normal sections
+
+\bibliographystyle{elsarticle-harv}
+\bibliography{main}
+
+\newpage
 \appendix
@@
 \begin{center}
-	\centering
-	\includegraphics[width=0.4\textwidth]{images/ocr_average_accuracy.png}	
+	\includegraphics[width=0.45\textwidth]{images/ocr_average_accuracy.png}	
 	\captionof{figure}{Average accuracy score for reading the sample dataset for the three OCR models}
@@
 \begin{center}
-	\centering
-	\includegraphics[width=0.4\textwidth]{images/pruned_ocr_average_accuracy.png}	
+	\includegraphics[width=0.45\textwidth]{images/pruned_ocr_average_accuracy.png}	
 	\captionof{figure}{Average accuracy score only including non-zero values (only images that were able to be read)}
@@
 \begin{center}
-	\centering
-	\includegraphics[width=0.4\textwidth]{images/ocr_time_taken.png}	
+	\includegraphics[width=0.45\textwidth]{images/ocr_time_taken.png}	
 	\captionof{figure}{Total time taken for the three OCR models on the sample dataset}
@@
 
-%% \label{}
+\section{Examples in the Leaf Segmentation Process}
+\label{app_leaf_seg}
 
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%https://www.overleaf.com/project/64bec4b7b47d12a55db700f5
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/leaf_seg_example.png}
+    \captionof{figure}{Example leaf segmentation image generated by the SAM/ONNX Runtime}
+\end{center}
 
-\bibliographystyle{elsarticle-harv}
-\bibliography{main}
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/leaf_example.png}
+    \captionof{figure}{Example leaf generated by cropping and rotating a filtered segmentation}
+\end{center}
 
-%%\begin{thebibliography}{00}
+\section{Examples of Images with Obstructed Labels}
+\label{app_blocked}
 
-%% \bibitem[Author(year)]{label}
-%% For example:
+\begin{center}
+    \includegraphics[width=0.40\textwidth]{images/blocked/blocked_leaf.jpg}
+    \captionof{figure}{Example of label blocked by a leaf}
+\end{center}
 
-%% \bibitem[Aladro et al.(2015)]{Aladro15} Aladro, R., Martín, S., Riquelme, D., et al. 2015, \aas, 579, A101
+\begin{center}
+    \includegraphics[width=0.40\textwidth]{images/blocked/blocked_hand.jpg}
+    \captionof{figure}{Example of label blocked by the researcher's hand}
+\end{center}
 
-%%\end{thebibliography}
+\begin{center}
+    \includegraphics[width=0.40\textwidth]{images/blocked/blocked_out.jpg}
+    \captionof{figure}{Example of label cut by the boundary of the image}
+\end{center}
 
 \end{document}
-
-\endinput
-%%
-%% End of file `elsarticle-template-harv.tex'.

### 2023-07-29 scripts/startup.sh
updates to stuff

@@ -1,7 +1,6 @@
 # run basic startup scripts already written
 . scripts/venv.sh
 ./scripts/load_files.sh
-./scripts/build_paper.sh
 
 # get the files required for step 2
 cd step2

### 2023-07-29 paper/main.tex
paper slightly changed

@@
 Optical character recognition (OCR) was used to read these labels on the plants, image segmentation techniques in conjunction with machine learning algorithms were used for morphological classifications, machine learning models were used to predict treatment based on those classifications, and analyzed encoded EXIF tags were used for the purpose of finding leaf size and correlations between phenotypes.
-We found that our OCR model successfully read 91.3\% of the labeled plants, allowing for the information to be accurately placed in a spreadsheet.
+We found that our OCR model had an accuracy of 94.31\% for non-null text extractions, allowing for the information to be accurately placed in a spreadsheet.
 Our classification models identified leaf shape, color, and level of brown splotches with an average accuracy of 62.82\%, and plant treatment with an accuracy of 60.08\%.

### 2023-07-29 paper/main.tex
updated everythin

@@
 
-\makeatletter
-\def\ps@pprintTitle{%
-  \let\@oddhead\@empty
-  \let\@evenhead\@empty
-  \let\@oddfoot\@empty
-  \let\@evenfoot\@oddfoot
-}
-\makeatother
+\begin{document}
 
-\begin{document}
+\journal{SMCDC23}
+\renewcommand*{\today}{July 28, 2023}
 
@@
 
-\author[]{Vivaan Singhvi}
-\author[]{Langalibalele Lunga}
-\author[]{Pragya Nidhi}
-\author[]{Chris Keum}
-\author[]{Varrun Prakash}
+\author[fhs]{Vivaan Singhvi}
+\author[fhs]{Langalibalele Lunga}
+\author[fhs]{Pragya Nidhi}
+\author[fhs]{Chris Keum}
+\author[fhs]{Varrun Prakash}
+
+\affiliation[fhs]{organization={Farragut High School},
+            addressline={11237 Kingston Pike},
+            city={Knoxville},
+            postcode={37922},
+            state={Tennesee},
+            country={United States}}
 
@@
 Using the EXIF tags, we were able to retrieve the coordinates where the image was taken.
-We noticed all of the images contained were taken in approximately the same two locations, so any weather or soil API would be unable to provide specific conditions.
+We noticed all of the images contained were taken in approximately the same location, so any weather or soil API would be unable to provide specific conditions.
 If we were given this information as part of the dataset, we would have been better able to find correlations.
@@
 
-We would like to thank Dr. Ty Frazier for his suggestions and mentoring throughout this project.
+We thank Dr. Ty Frazier at Oak Ridge National Laboratory for his helpful suggestions and mentoring throughout this project.
 

### 2023-07-29 paper/main.tex
misspelled Tennessee.. ... . .. ..

@@
             postcode={37922},
-            state={Tennesee},
+            state={Tennessee},
             country={United States}}
@@
 \begin{abstract}
-%% Text of abstract
 High-throughput phenotyping refers to the non-destructive and efficient evaluation of plant phenotypes.
@@
 \end{frontmatter}
-
-%\tableofcontents
-
-%% \linenumbers
-
-%% main text
 
@@
 
-To read the text, a pre-trained Optical Character Recognition (OCR) model seemed like the optimal choice.
+To read the text, a pre-trained optical character recognition (OCR) model seemed like the optimal choice.
 However, there are several high-performing models available for use.
@@
 With the world population projected to reach 9.3 billion people by 2050 and a need to produce 60\% more food \citep{silva2012feeding}, it will become crucial for researchers to discover a method for identifying productive genotypes for plant breeding.
-With computer vision techniques, such as Optical Character Recognition and image processing, and machine learning methods, such as classifiers to predict morphologies and treatments, productivity in the field of plant/crop production will increase rapidly, along with the accurate analysis of plants.
+With computer vision techniques, such as optical character recognition and image processing, and machine learning methods, such as classifiers to predict morphologies and treatments, productivity in the field of plant/crop production will increase rapidly, along with the accurate analysis of plants.
 
@@
 
-This project is one of the first to utilize \verb|PaddleOCR| \citep{du2020ppocr}, an extremely lightweight and capable Optical Character Recognition model, in the context of reading plant labels.
+This project is one of the first to utilize \verb|PaddleOCR| \citep{du2020ppocr}, an extremely lightweight and capable optical character recognition model, in the context of reading plant labels.
 Unlike other approaches, it is resilient to augmentations in data, such as label rotation or minor interruptions in labels.
@@
 
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-
 \bibliographystyle{elsarticle-harv}
@@
 	\captionof{figure}{Average accuracy score for reading the sample dataset for the three OCR models}
-	% \label{fig_mom0}%
 \end{center}
@@
 	\captionof{figure}{Average accuracy score only including non-zero values (only images that were able to be read)}
-	% \label{fig_mom0}%
 \end{center}
@@
 	\captionof{figure}{Total time taken for the three OCR models on the sample dataset}
-	% \label{fig_mom0}%
 \end{center}

### 2023-08-16 paper/poster.tex
added poster pdf and texfile

\documentclass[final]{beamer}

\usepackage[size=custom, width=36, height=24, scale=0.36]{beamerposter}
\usetheme{gemini}
\usecolortheme{msu}
\usepackage[labelfont=bf]{caption}
\usepackage{picins}

\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.025\paperwidth}
\setlength{\colwidth}{0.3\paperwidth}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

\title{High-Throughput Phenotyping using Computer Vision and Machine Learning}

\author{Vivaan Singhvi, Langalibalele Lunga, Pragya Nidhi, Chris Keum, Varrun Prakash}

\institute{Farragut High School}

\footercontent{
    \href{https://github.com/vivaansinghvi07/smoky-mountain-data-comp}{Github: vivaansinghvi07/smoky-mountain-data-comp} 
    \hfill
    Smoky Mountains Computational Sciences Data Challenge Conference 
    \hfill
    \href{mailto:singhvi.vivaan@gmail.com}{singhvi.vivaan@gmail.com}
}

\begin{document}

\begin{frame}[t]
\begin{columns}[t]
\separatorcolumn

\begin{column}{\colwidth}

\begin{block}{Introduction and Background Information}

% ---------------------------------------------------- %

\textbf{High-throughput phenotyping} is a breakthrough technology used in plant biology and agriculture to examine a wide variety of plant features through \textbf{image analysis}. By coupling this technique with \textbf{machine learning}, scientists can handle plant datasets \textbf{thousands of times larger} in a \textbf{fraction of the time}.

% ---------------------------------------------------- %

\end{block}

\begin{alertblock}{Research Objective}

% ---------------------------------------------------- %

Using a dataset of \textbf{1672} images (with spanning \textbf{white labels}) of the plant \textbf{Populus Trichocarpa}, provided by Oak Ridge National Laboratory, we aim to address the following summarized challenge questions:

% ---------------------------------------------------- %

\begin{enumerate}

    \item Is it possible to use \textbf{optical character recognition} to “read” each label and \textbf{generate a spreadsheet} of the features on the label?
    \item Can \textbf{machine learning} classify different \textbf{leaf morphologies} among plants, such as leaf shape or color?
    \item Can a \textbf{predictive model} be built \textbf{using leaf morphology classifications} that can indicate the \textbf{condition} in which a plant was raised?
    \item GPS and other camera information are encoded in \textbf{EXIF tags}. Can this data be used to determine characteristics such as leaf size? Can other data, such as soil maps, weather, etc. be used to find \textbf{correlations among phenotypes}?

\end{enumerate}

% ---------------------------------------------------- %

\end{alertblock}

\begin{block}{Reading Labels with Optical Character Recognition}

% ---------------------------------------------------- %

\begin{itemize}
    \item The prebuilt model \textbf{PaddleOCR} was chosen for label reading due to its \textbf{accuracy} and \textbf{efficiency}.
    \item However, minor \textbf{image augmentation} was needed to configure images, including \textbf{edge highlighting} and \textbf{rotating}.
\end{itemize}

% ---------------------------------------------------- %

\begin{center}
    \includegraphics[width=0.87\colwidth]{images/step1_pipeline.png}
    \captionof{figure}{The image augmentation process for optical character recognition}
\end{center}

\vspace{-0.5\intextsep}

% ---------------------------------------------------- %

\begin{itemize}
    \parpic(0px, 0px)(0px, 1.7cm)[r]{
        \resizebox{0.57\colwidth}{!}{
            \begin{minipage}{\colwidth}
                \captionsetup{font={Large}}
                \begin{tabular}{|c|c|c|c|c|c}
                     \hline
                     filename & treatment & block & row & position & genotype \\
                     \hline
                        ... & D & 1 & 8  & 32 & BESC-34 \\
                        ... & C & 1 & 10 & 12 & **BESC-417\_LM**,core \\
                        ... & C & 2 & 3  & 40 & BESC-468 \\
                        ... & C & 2 & 6  & 54 & BESC-28\_LM \\
                        ... & C & 1 & 24 & 22 & **LILD-26-5\_LM**,core \\
                        ... & 	& 	& 	 &	  & **HOMD-21-2\_LM**,core \\
                        ... & C	& 2	& 23 & 45 & BESC-361\_16\_11\_CB \\
            	      ... &   & 1 & 25 & 30	& **BESC-106\_LM**,core \\
                     \hline
                \end{tabular}
                \captionof{table}{Example spreadsheet data}
            \end{minipage}
        }
    }
    \item \textbf{Regular Expressions} extracted features from text.
    \item On 30 random images, the model had an accuracy of \textbf{77.33\%}, \textbf{94.31\%} with null values omitted.
\end{itemize}

% ---------------------------------------------------- %

\end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

\begin{block}{Classifying Leaf Morphologies with Image Segmentation}

% ---------------------------------------------------- %

\begin{itemize}
    \item The \textbf{Segment Anything Model} (SAM) was chosen to segment leaves. The \texttt{SamPredictor} could generate an exact mask at a \textbf{given point}.
    \item To approximate these points, the below pipeline was used, using \textbf{HSV Filtering}, \textbf{Edge Detection}, and \textbf{Dilation} techniques.
\end{itemize}

% ---------------------------------------------------- %

\vspace{-1em}

\begin{center}
    \includegraphics[width=0.22\colwidth]{1.png}
    \includegraphics[width=0.22\colwidth]{2.png}
    \includegraphics[width=0.22\colwidth]{4.png}
    \includegraphics[width=0.22\colwidth]{6.png}
    \captionof{figure}{The image processing pipeline utilized to generate leaf approximations}
\end{center}

\vspace{-0.5\intextsep}

% ---------------------------------------------------- %

\vspace{-0.7em}

\begin{itemize}
    \parpic(0px, 0px)(0px, 2.75cm)[l]{
        \begin{minipage}{0.28\colwidth}
            \includegraphics[width=0.28\colwidth]{images/2_better.png}
            \captionof{figure}{Segmentation masks after filtering}
        \end{minipage}
    }
    \item The \textbf{centers} of these approximations were used as the \textbf{generation point}.
    \item Then, using a \textbf{machine learning classifier}, we \textbf{filtered bad segmentations} (shown in red to the left) with an accuracy of \textbf{90.91\%}.
    \item The masks are then used to \textbf{classify morphologies}.
    \item The chosen features were \textbf{color}, \textbf{shape}, and \textbf{level of brown splotch} (indicating withering leaves).
    \item The \texttt{XGBoost} model classified the features, and \texttt{scikit-learn}'s \texttt{MultiOutputClassifier} was implemented to manage all three at once.
\end{itemize}

% ---------------------------------------------------- %

\vspace{-0.5em}

\begin{center}
    \includegraphics[width=0.32\colwidth]{images/color.png}
    \includegraphics[width=0.32\colwidth]{images/shape.png}
    \includegraphics[width=0.32\colwidth]{images/splotch.png}
    \captionof{figure}{Confusion matrices for color, shape, and splotch respectively}
\end{center}

% ---------------------------------------------------- %

\vspace{-1em}

\begin{itemize}
    \item The \textbf{color} classifier had an accuracy of \textbf{69.23\%}, and seemed to be confused around \textbf{true yellow-green} leaves.
    \item The \textbf{shape} classifier was more sporadic with an accuracy of \textbf{50.00\%}, \textbf{mispredicting} \textbf{oblong} and \textbf{ovate} leaves the most.
    \item The \textbf{splotch} classifier had an accuracy of \textbf{69.23\%}, without significant errors.
    \item The model was used on \textbf{every image} by finding the \textbf{mode prediction} for each feature. Then, this data was \textbf{inserted into the spreadsheet}. 
\end{itemize}

% ---------------------------------------------------- %

\end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

\begin{block}{Predicting Treatment from Morphological Classifications}

% ---------------------------------------------------- %

\begin{itemize}
    \item By using \textbf{read treatments} from step 1 in conjunction with \textbf{morphological classifications} from step 2, we could build a simple predictor to determine if a plant was raised in \textbf{drought} or \textbf{control}.
\end{itemize}

\vspace{-1em}

% ---------------------------------------------------- %

\begin{itemize}
    \parpic(0px, 0px)(0px, 1.3cm)[r]{
        \resizebox{0.57\colwidth}{!}{
            \begin{minipage}{\colwidth}
                \begin{tabular}{|c||c c c c|} 
                     \hline
                     leaf\_color & light\_green & dark\_green & yellow\_green & yellow  \\
                     \hline
                     light\_green & 1 & 0 & 0 & 0 \\
                     yellow\_green & 0 & 0 & 1 & 0 \\
                     dark\_green & 0 & 1 & 0 & 0 \\ 
                     light\_green & 1 & 0 & 0 & 0 \\
                     yellow & 0 & 0 & 0 & 1 \\
                     \hline
                \end{tabular}
                \captionsetup{font={Large}}
                \captionof{table}{Example of one-hot encoding}
            \end{minipage}
        }
    }
    \item \textbf{One-hot encoding}, as seen to the right, was used to convert our \textbf{qualitative} data to \textbf{quantitative} data.
\end{itemize}

% ---------------------------------------------------- %

\begin{itemize}
    \parpic(0px, 0px)(-0.2cm, 2.6cm)[l]{
        \begin{minipage}{0.35\colwidth}
            \includegraphics[width=0.35\colwidth]{images/treat_conf_matrix.png}
            \captionof{figure}{Confusion matrix for treatment classifier}
        \end{minipage}
    }
    \item Some data had to be \textbf{pruned} as to avoid \textbf{class imbalance}.
    \item Our model, a \texttt{RandomForestClassifier}, had an accuracy of \textbf{60.08\%} and a confusion matrix shown on the left.
    \item The low accuracy is likely due to our limitation of \textbf{only using classification data} from the previous step. Along with the fact that they \textbf{may be inaccurate}, only three features are \textbf{likely not enough} to make good predictions.
\end{itemize}

% ---------------------------------------------------- %

\end{block}

\vspace{1em}

\begin{block}{Finding Correlations and Characteristics from EXIF Tags}

% ---------------------------------------------------- %

\begin{itemize}
    \item The EXIF tags were \textbf{not useable} for predicting leaf size, or other traits. 
    \item A vital tag, the \texttt{FocalPlaneResolution}, was \textbf{missing} from the images.
    \item Additionally, since all leaves were close together, \textbf{no weather or soil map API} would provide geolocational data \textbf{specific enough} for us.
    \item If this information were \textbf{present in a given file}, we could make conclusions.
\end{itemize}

% ---------------------------------------------------- %

\end{block}

\begin{exampleblock}{Conclusion and Significance}

% ---------------------------------------------------- %

\begin{itemize}
    \item We were \textbf{successfully} able to read plant labels with \textbf{optical character recognition} and store the data in our spreadsheet. 
    \item We were \textbf{successfully} able to extract leaves from the images and \textbf{somewhat successfully} able to classify their \textbf{morphologies}.
    \item We were \textbf{slightly successfully} able to determine \textbf{treatment} from morphology classifications and \textbf{not successful} in making conclusions from \textbf{EXIF tags}.
    \item Regarding \textbf{originality}, this study is \textbf{one of the first} to implement \textbf{PaddleOCR} and the \textbf{Segment Anything Model} in this context, due to their relative recency. Their powerful capabilities proved to be \textbf{vital} to our research.
\end{itemize}

% ---------------------------------------------------- %

\end{exampleblock}

\end{column}

\separatorcolumn

\end{columns}

\end{frame}

\end{document}

### 2023-10-04 paper/main.tex
better display?
built new verison of paper

@@
 Traditionally, plant morphologies would be analyzed manually, wasting time and potentially damaging the environment.
-However, current breakthroughs in high-throughput phenotyping led by machine learning techniques\citep{koh2021automated, pound2017deep} and computer vision with deep learning \citep{mochida2019computer} have allowed for efficient and timely analyses of plant phenotypes.
+However, current breakthroughs in high-throughput phenotyping led by machine learning techniques \citep{koh2021automated, pound2017deep} and computer vision with deep learning \citep{mochida2019computer} have allowed for efficient and timely analyses of plant phenotypes.
 With these methods, researchers can extract features from plants and classify their morphologies effectively while having little effect on the plant itself.
@@
 Plants are now able to be analyzed in fast, effective, and accurate methods allowing for them to be used in scientific discoveries.
-In our study, we rely on machine learning techniques to classify morphologies based on a phone image dataset provided.
+In our study, we rely on machine learning techniques to classify morphologies based on a plant image dataset provided.
 
@@
 	\includegraphics[width=0.23\textwidth]{images/seg_pipeline/6.png}	
-        \includegraphics[width=0.23\textwidth]{images/seg_filter/2.png}	
+        \includegraphics[width=0.23\textwidth]{images/seg_filter/2_better.png}	
 	\captionof{figure}{From leaf approximations with image processing to filtered leaf segmentations, from the SAM/ONNX Runtime (green leaves fit for classification)}
@@
 Using the EXIF tags, we were able to retrieve the coordinates where the image was taken.
-We noticed all of the images contained were taken in approximately the same location, so any weather or soil API would be unable to provide specific conditions.
+However, we noticed all of the images contained were taken in approximately the same location.
+
+\begin{center}
+    \includegraphics[width=0.45\textwidth]{images/plant_locs.png}
+    \captionof{figure}{Geographical plot of all plants using encoded EXIF GPS data.}
+\end{center}
+
+Therefore, any weather or soil API would be unable to provide specific conditions.
 If we were given this information as part of the dataset, we would have been better able to find correlations.
@@
 
-Future research could be done in expanding the species analyzed for morphology classifications.
+Future research could be done to expand the species analyzed for morphology classifications.
 In this study, only Populus Trichocarpae were used, so the models used may only be effective on this species.
 
-Since a pretrained segmentation model (the SAM) was used in this study, researchers could attempt to build segmentation models fine-tuned to only recognize leaves, which could increase model efficiency and provide more consistent results.
+Since a pre-trained segmentation model (the SAM) was used in this study, researchers could attempt to build segmentation models fine-tuned to only recognize leaves, which could increase model efficiency and provide more consistent results.
 
@@
 
-\newpage
+\clearpage
 \appendix
